Exploring Income Inequality Dynamics in Canada: A Geospatial and Temporal Analysis

### 2024-04-03 paper/paper.qmd
change

begin{align}
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \
\mu_i &= \beta_0 + \beta_1 \times \text{Year}_i + \beta_2 \times \text{Geographical Location}_i\
\beta_0 &\sim \text{Normal}(0, 2.5) \
\beta_1 &\sim \text{Normal}(0, 2.5) \
\beta_2 &\sim \text{Normal}(0, 2.5) \
\sigma &\sim \text{Exponential}(1)
\label{eq-Model2}
\end{align}

In Model (\ref{eq-Model1})is specified as follows:

-   $Y_{ij}$ represents the highest-to-Lowest average income ratio of average after-tax income to the lowest decile of average after-tax income of Canada in $i^{th}$ year and province $j$.

@@

y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \
\mu_i &= \beta_0 + \beta_1 \times \text{Year}_i + \beta_2 \times \text{Geographical Location}_i\
\beta_0 &\sim \text{Normal}(0, 2.5) \
\beta_1 &\sim \text{Normal}(0, 2.5) \
\beta_2 &\sim \text{Normal}(0, 2.5) \
\sigma &\sim \text{Exponential}(1)
\label{eq-Model2}
\end{align}

In Model (\ref{eq-Model2}):

-   $Y_{ij}$ represents the income range for a particular combination of year ($Year_{i}$) and geographical location ($Province_{j}$).

@@

Model (\ref{eq-Model2}) is formulated to explore the relationship between the income range ($Y_{ij}$) and the predictor variables, Year ($Year_{i}$), and Geographical Location ($Province_{j}$).

$Y_{ij}$ represents the income range for a specific combination of year ($Year_{i}$) and geographical location ($Province_{j}$). This formulation allows us to examine how income ranges vary across different years and provinces.

@@

$\beta_{2}$ is the coefficient associated with the "Geographical Location" variable ($Province_{j}$). This coefficient captures the average difference in income range between different provinces, assuming the year remains constant. It allows us to assess the impact of geographical location on income disparities.

Model (\ref{eq-Model2}) provides a structured framework for analyzing the factors influencing income range variations across different years and geographical locations. By examining the effects of both temporal and spatial factors, the model helps to elucidate patterns of income distribution and disparities within the studied context.

@@

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false

library(rstanarm)

@@

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

second_model<- 
  readRDS(file = here::here("models/model.rds"))
```

```{r}
#| label: tbl-model1
#| tbl-cap: "Summary results for first model"
#| echo: false
#| eval: true
#| warning: false
#| message: false
library(rstanarm)

@@

@tbl-modelresults2

The results of the model reveal several significant findings regarding the relationship between the response variable and the predictor variables. Firstly, the intercept term is estimated to be -3309430.66, indicating the expected value of the response variable when all other predictors are zero. This intercept provides a reference point for comparison and interpretation of the effects of the other predictor variables.

>>>>

\begin{align}
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{Year}_i + \beta_2 \times \mbox{Geographical Location}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\label{eq-model1}
\end{align}

In Model (\ref{eq-model1})is specified as follows:

-   $Y_{i}$ represents the highest-to-Lowest average income ratio of average after-tax income to the lowest decile of average after-tax income of Canada in $i^{th}$ year and province $i$.

@@

\begin{align}
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \text{Year}_i + \beta_2 \times \text{Geographical Location}_i\\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\beta_1 &\sim \text{Normal}(0, 2.5) \\
\beta_2 &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}(1)
\label{eq-model2}
\end{align}

In Model (\ref{eq-model2}):

-   $Y_{i}$ represents the income range for a particular combination of year ($Year_{i}$) and geographical location ($Province_{j}$).

@@

Model (\ref{eq-model2}) is formulated to explore the relationship between the income range ($Y_{ij}$) and the predictor variables, Year ($Year_{i}$), and Geographical Location ($Province_{j}$).

$Y_{ij}$ represents the income range for a specific combination of year ($Year_{i}$) and geographical location ($Province_{i}$). This formulation allows us to examine how income ranges vary across different years and provinces.

@@

$\beta_{2}$ is the coefficient associated with the "Geographical Location" variable ($Province_{i}$). This coefficient captures the average difference in income range between different provinces, assuming the year remains constant. It allows us to assess the impact of geographical location on income disparities.

Model 2 provides a structured framework for analyzing the factors influencing income range variations across different years and geographical locations. By examining the effects of both temporal and spatial factors, the model helps to elucidate patterns of income distribution and disparities within the studied context.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
library(rstanarm)

@@

modelsummary(
  list(
    "Model 1" = first_model,
    "Model 2" = second_model
  ), statistic = "mad",
  fmt = 2
)

@@

```{r}
#| label: tbl-model1
#| tbl-cap: "Summary results for two models"
#| echo: false
#| eval: true
#| warning: false
#| message: false
first_model <- 
  readRDS(file = here::here("models/model2.rds"))

@@

The results of the model2 reveal several significant findings regarding the relationship between the response variable and the predictor variables. Firstly, the intercept term is estimated to be -3309430.66, indicating the expected value of the response variable when all other predictors are zero. This intercept provides a reference point for comparison and interpretation of the effects of the other predictor variables.
